---
title: "CV of Plastic and Immutable Scenario with different S.D. "
format: pdf
editor: visual
---

```{r}
pacman::p_load(ggplot2,#plots
               tictoc, #timing
               dplyr, #filter 
               e1071 #skewness,
)
```

```{r}
source("../functions_main/other_functions.R")
```

Space

Space

Space

#### CV analysis

I computed separately on the server 500 simulation for each case:

|      |             |                   |          |                          |
|------|-------------|-------------------|----------|--------------------------|
| Case | Trait given | Coefficient Trait | Type     | Standard deviation Trait |
| 1    | Once        | 1                 | None     | 0                        |
| 2    | Once        | 1                 | Small    | 0.15                     |
| 3    | Once        | 1                 | Moderate | 0.3                      |
| 4    | Once        | 1                 | High     | 0.45                     |

|      |             |                   |          |                          |
|------|-------------|-------------------|----------|--------------------------|
| Case | Trait given | Coefficient Trait | Type     | Standard deviation Trait |
| 5    | Every year  | 1                 | None     | 0                        |
| 6    | Every year  | 1                 | Small    | 0.15                     |
| 7    | Every year  | 1                 | Moderate | 0.3                      |
| 8    | Every year  | 1                 | High     | 0.45                     |

|  |  |  |  |  |  |
|----|----|----|----|----|----|
| Case | Trait given | Coefficient Trait | Type | SD Trait | Coefficient of plasticity |
| 9 | Updated partially every year | 1 | None | 0 | 0.5 |
| 10 | Updated partially every year | 1 | Small  | 0.15 | 0.5 |
| 11 | Updated partially every year | 1 | Moderate | 0.3 | 0.5 |
| 12 | Updated partially every year | 1 | High | 0.45 | 0.5 |

### Immutable scenario

```{r}
case9 <-readRDS("../server_runs/RDS/case9.rds")
case10<-readRDS("../server_runs/RDS/case10.rds")
case11<- readRDS("../server_runs/RDS/case11.rds")
case12 <- readRDS("../server_runs/RDS/case12.rds")
```

```{r}
mix_plasticity_cv <- bind_rows(
  data.frame(cv = coefficient_of_variation(case9), sd = "none"),
  data.frame(cv = coefficient_of_variation(case10), sd = "low"),
  data.frame(cv = coefficient_of_variation(case11), sd = "medium"),
  data.frame(cv = coefficient_of_variation(case12), sd = "high")) |>
  mutate(sd = factor(sd, levels = c("none","low", "medium", "high")))
```

```{r}
mycolors <- c("#728f7c","#5da174","#3dba67", "#04cf48")
p1 <- ggplot(mix_plasticity_cv, aes(x = cv, fill = sd)) +
  geom_histogram(binwidth = 0.13, color = "black", alpha = 0.6) +
  geom_vline(
    data = mix_plasticity_cv |> group_by(sd) |> summarise(mean_cv = mean(cv)),
    aes(xintercept = mean_cv),
    color = "black", linetype = "dashed", linewidth = 1
  ) +
  labs(x = "Coefficient of Variation", y = "Counts of simulations") +
  theme_minimal() +
  coord_flip() +
  xlim(0,7)+
  scale_y_continuous(breaks = c(30, 60))+
  facet_wrap(~sd, ncol = 4, strip.position = "top") +
  scale_fill_manual(values = mycolors) +
  ggtitle("Mixed Plasticity Scenario") +
  theme(
    legend.position = "none",
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x = element_text(size = 14),   
    axis.text.y = element_text(size = 14),  
    plot.title = element_text(size = 20, face = "bold", hjust=0.5),
    strip.text = element_text(size = 16)
  )
ggsave("mix_plasticity_cv.png", plot = p1, width = 5, height = 6, dpi = 300)
p1
```

```{r}
mix_plasticity_cv<- mix_plasticity_cv |>
       group_by(sd) |>
            summarise(mean_cv = mean(cv))
mix_plasticity_cv
```

```{r}
mix_plasticity_extinction <- c(extinct_count(case9),extinct_count(case10),extinct_count(case11), extinct_count(case12))
sd_labels <- c("none","low", "medium", "high")

mix_plasticity_extinction_df <- data.frame(
  sd = factor(sd_labels, levels = sd_labels),
  count = mix_plasticity_extinction
)


mix_plasticity_extinction_df
```

```{r}

p2 <- ggplot(mix_plasticity_extinction_df, aes(x = sd, y = count, fill = sd)) +
  geom_bar(stat = "identity", color = "black", alpha = 0.7) +
  scale_fill_manual(values = mycolors) +
  labs(
    x = "SD Category",
    y = "Extinction Count",
    title = "Mixed Plasticity Scenario"
  ) +
  ylim(0,420)+
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5)
  )


ggsave("extinction_count_mixed_scenario.png", plot = p2, width = 5, height = 6, dpi = 300)
p2
```

Now we have to compare the three scenarios
