---
title: "CV of Plastic and Immutable Scenario with different S.D. "
format: pdf
editor: visual
---

```{r}
pacman::p_load(ggplot2,#plots
               tictoc, #timing
               dplyr, #filter 
               e1071 #skewness,
)
```

```{r}
source("../functions_main/other_functions.R")
```

Space

Space

Space

#### CV analysis

I computed separately on the server 500 simulation for each case:

|      |             |                   |          |                          |
|------|-------------|-------------------|----------|--------------------------|
| Case | Trait given | Coefficient Trait | Type     | Standard deviation Trait |
| 1    | Once        | 1                 | None     | 0                        |
| 2    | Once        | 1                 | Small    | 0.15                     |
| 3    | Once        | 1                 | Moderate | 0.3                      |
| 4    | Once        | 1                 | High     | 0.45                     |

|      |             |                   |          |                          |
|------|-------------|-------------------|----------|--------------------------|
| Case | Trait given | Coefficient Trait | Type     | Standard deviation Trait |
| 5    | Every year  | 1                 | None     | 0                        |
| 6    | Every year  | 1                 | Small    | 0.15                     |
| 7    | Every year  | 1                 | Moderate | 0.3                      |
| 8    | Every year  | 1                 | High     | 0.45                     |

### Immutable scenario

```{r}
case1 <-readRDS("../job_sim/RDS/case1.rds")
case2<-readRDS("../job_sim/RDS/case2.rds")
case3<- readRDS("../job_sim/RDS/case3.rds")
case4 <- readRDS("../job_sim/RDS/case4.rds")
```

```{r}
immutable_cv <- bind_rows(
  data.frame(cv = coefficient_of_variation(case1), sd = "none"),
  data.frame(cv = coefficient_of_variation(case2), sd = "low"),
  data.frame(cv = coefficient_of_variation(case3), sd = "medium"),
  data.frame(cv = coefficient_of_variation(case4), sd = "high")) |>
  mutate(sd = factor(sd, levels = c("none","low", "medium", "high")))
```

```{r}
mycolors <- c("#edeef5","#c5caf0","#7083fa", "#0317ad")
p1 <- ggplot(immutable_cv, aes(x = cv, fill = sd)) +
  geom_histogram(binwidth = 0.13, color = "black", alpha = 0.6) +
  geom_vline(
    data = immutable_cv |> group_by(sd) |> summarise(mean_cv = mean(cv)),
    aes(xintercept = mean_cv),
    color = "black", linetype = "dashed", linewidth = 1
  ) +
  labs(x = "Coefficient of Variation", y = "Counts of simulations") +
  theme_minimal() +
  coord_flip() +
  xlim(0,7)+
  scale_y_continuous(breaks = c(30, 60))+
  facet_wrap(~sd, ncol = 4, strip.position = "top") +
  scale_fill_manual(values = mycolors) +
  ggtitle("Early Life Plasticity Scenario") +
  theme(
    legend.position = "none",
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x = element_text(size = 14),   # numeri asse x più grandi
    axis.text.y = element_text(size = 14),   # numeri asse y più grandi
    plot.title = element_text(size = 20, face = "bold", hjust=0.5),
    strip.text = element_text(size = 16)
  )
p1
ggsave("plot1.png", plot = p1, width = 5, height = 6, dpi = 300)
```

```{r}
immutable_cv_table<- immutable_cv |>
       group_by(sd) |>
            summarise(mean_cv = mean(cv), sd =sd(cv))
immutable_cv_table
```

```{r}
immutable_extinction <- c(extinct_count(case1),extinct_count(case2),extinct_count(case3), extinct_count(case4))
sd_labels <- c("none","low", "medium", "high")

immutable_extinction_df <- data.frame(
  count = immutable_extinction/500,
  sd = factor(sd_labels, levels = sd_labels)
)

```

```{r}

p2 <- ggplot(immutable_extinction_df, aes(x = sd, y = count, fill = sd)) +
  geom_bar(stat = "identity", color = "black", alpha = 0.7) +
  scale_fill_manual(values = mycolors) +
  labs(
    x = "SD Category",
    y = "Probability of Extinction",
    title = "Early Life Plasticity Scenario"
  ) +
  ylim(0,1)+
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5)
  )

p2
ggsave("plot2.png", plot = p2, width = 5, height = 6, dpi = 300)
```

#### Plasticity scenario

Every year the traits of all the alive animals are updated

```{r}
case5 <-readRDS("../job_sim/RDS/case5.rds")
case6<-readRDS("../job_sim/RDS/case6.rds")
case7<- readRDS("../job_sim/RDS/case7.rds")
case8 <- readRDS("../job_sim/RDS/case8.rds")
```

```{r}
plasticity_cv <- bind_rows(
  data.frame(cv = coefficient_of_variation(case5), sd = "none"),
  data.frame(cv = coefficient_of_variation(case6), sd = "low"),
  data.frame(cv = coefficient_of_variation(case7), sd = "medium"),
  data.frame(cv = coefficient_of_variation(case8), sd = "high")) |>
  mutate(sd = factor(sd, levels = c("none","low", "medium", "high")))
```

```{r}
mycolors <- c("#faf5f6","#e0abb4","#ad3b50","#5e0112")
p3 <- ggplot(plasticity_cv, aes(x = cv, fill = sd)) +
  geom_histogram(binwidth = 0.13, color = "black", alpha = 0.6) +
  geom_vline(
    data = plasticity_cv |> group_by(sd) |> summarise(mean_cv = mean(cv)),
    aes(xintercept = mean_cv),
    color = "black", linetype = "dashed", linewidth = 1
  ) +
  labs(x = "Coefficient of Variation", y = "Counts of simulations") +
  theme_minimal() +
  coord_flip() +
  xlim(0,7)+
  scale_y_continuous(breaks = c(30, 60))+
  facet_wrap(~sd, ncol = 4, strip.position = "top") +
  scale_fill_manual(values = mycolors) +
  ggtitle("Full Plasticity Scenario") +
  theme(
    legend.position = "none",
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x = element_text(size = 14),  
    axis.text.y = element_text(size = 14),  
    plot.title = element_text(size = 20, face = "bold", hjust=0.5),
    strip.text = element_text(size = 16)
  )
p3
ggsave("plot3.png", plot = p3, width = 5, height = 6, dpi = 300)
```

```{r}
plasticity_cv_table<- plasticity_cv |>
       group_by(sd) |>
            summarise(mean_cv = mean(cv),sd= sd(cv))
plasticity_cv_table
```

```{r}
plasticity_extinction <- c(extinct_count(case5),extinct_count(case6),extinct_count(case7), extinct_count(case8))
sd_labels <- c("none","low", "medium", "high")

plasticity_extinction_df <- data.frame(
  count = plasticity_extinction/500,
  sd = factor(sd_labels, levels = sd_labels)
)

```

```{r}
p4 <- ggplot(plasticity_extinction_df, aes(x = sd, y = count, fill = sd)) +
  geom_bar(stat = "identity", color = "black", alpha = 0.7) +
  scale_fill_manual(values = mycolors) +
  labs(
    x = "SD Category",
    y = "Extinction Count",
    title = "Full Plasticity Scenario"
  ) +
  ylim(0,1)+
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5)
  )

p4
ggsave("plot4.png", plot = p4, width = 5, height = 6, dpi = 300) #maybe put it back to 6 and 4 later on
```

The count of extinctions is hugely greater than the trait given only at birth scenario, why???

```{r}
data <- data.frame(
  SD = c("None", "Low", "Medium", "High"),
  `Mean CV Immutable`      = immutable_cv_table$mean_cv,
  `Mean CV Flexible`       = plasticity_cv_table$mean_cv,
  `Extinction % Immutable` = immutable_extinction_df$count ,
  `Extinction % Flexible`  = plasticity_extinction_df$count 
)

```

| Sd | Mean CV immutable | Mean CV Flexible | Extinct % Immutable | Extinct % Flexible |
|----|----|----|----|----|
| None | 1.123980 | 1.630573 | 0.216 | 0.810 |
| Low | 1.182187 | 1.639052 | 0.312 | 0.776 |
| Medium | 1.251113 | 1.679113 | 0.358 | 0.758 |
| High | 1.439831 | 1.839916 | 0.484 | 0.800 |

```{r}


n <- 1e6


# Environment
E <- rnorm(n, mean = 0, sd = 2)

# Case 2: fixed trait
T_fix <- 2
psurv_fix <- plogis(T_fix * E)

# Case 1: variable trait
T_var <- rnorm(n, mean = 2, sd = 0.6)
psurv_var <- plogis(T_var * E)


# mean(psurv_fix)
# mean(psurv_var)
mean(psurv_fix)
mean(psurv_var)


```
